
\documentclass[12 pt]{article}
\usepackage{fullpage}
\usepackage{rotating}
\usepackage{tipa}
\usepackage{amsmath}
\usepackage{linguex}
\usepackage{enumerate}
\usepackage[sc]{mathpazo}
\linespread{1.05}         % Palatino needs more leading (space between lines)
\usepackage[T1]{fontenc}

\hyphenpenalty= 7000 

\title{White Paper \\ The FieldLinguists' App}
\author{}
\date{}

\begin{document}
\maketitle{} 

%\tableofcontents{section}

\section {Project Abstract}

The FieldLinguists' App is an OpenSource database that will allow language researchers to securely enter, store, organize, annotate, and share linguistic data. The application will be accessible everywhere; it will run on three different systems (Mac, Linux, and Windows) as a Chrome extension and will be suitable for both online and offline use.  Furthermore, the application will be created with collaborative goals in mind;  data will be syncable and sharable with other researchers.  Researchers can form teams that contribute to a single corpus, where team members can modify and discuss the data from within the application. The system will also have a simple and friendly user interface, allowing participants to drag and drop files (audio, video, text), or record them directly into the database when using the Android app. In addition, the application will have import and export capabilities for multiple file types.  Most importantly, the application is designed intuitively and theory free, so it is not necessary to be a field linguist or programmer to figure out how it works.

%Data will be stored in a host server for sync/integration and long-term archival purposes. The app has a local storage space allowing quick search and offline work. 

\section {Statement of Need}

The FieldLinguists' App is conceived out of the needs of language researchers doing fieldwork or lather large scale data collection. Linguistic fieldwork often requires researchers to travel to places where a stable connection to the internet is not guaranteed. Also, it often involves a group of researchers contributing to building a single database.
 An ideal linguistic database should therefore work both on- and off-line as well as make it easy to share and integrate data. 
 
There are existing programmes/software used for linguistic fieldwork, 
however, they fall short in providing features necessary for collaborative fieldwork. 
For example, some web-based databases (e.g.  {\it Karuk Dictionary and Texts} http://linguistics.berkeley.edu/{\textasciitilde}ka
ruk/links.php,   {\it The Washo Project} http://washo.uchicago.edu/dictionary/dictionary.php) work only online, hence it is impossible for researchers to enter new data or search the database while in the field where the internet is unavailable.  

Non-web-based software such as {\it Toolbox} (http://www.sil.org/computing/toolbox/) and {\it FLEx/FieldWorks} (http://fieldworks.sil.org/flex/)
are excellent in terms of annotating data and organizing data into various formats (corpus, grammar or lexicon). Nonetheless, integration of data taken by individual researchers is not well-automated and takes considerable work by researchers. Moreover, they run only on a single platform (either PC or Linux, but not both). These tools therefore cause further difficulty in data sharing among researchers using different platforms.  

General purpose database software such as {\it FileMaker Pro} can be customized for the purpose of language research. However, it demands researchers to learn the software, and research teams often need to hire a programmer to customize the software for their research purposes. The stand-alone nature of such software makes data sharing and integration difficult as well. 

The existing linguistic database programmes, although useful, have various shortfalls that would hinder collection and integration of data. Some of them are constrained by the internet accessibility or by the computer platform types. Some others demand extra human work in order to integrate data. 
%A database application whose functionality is not limited by such constraints is much anticipated by language researchers. 
The present project grows out of discussion with a numbers of fieldworkers dissatisfied with currently available options. 


\section {App Description}

The FieldLinguists' App will enable those interested in language research, preservation, and documentation to securely enter, store, organize, annotate, and share linguistic data. Moreover, the application will be easily customizable to fit specific needs. To accomplish these tasks, the database will be equipped with a variety of features. 
The following functional requirements are based on a few important considerations where most existing fieldlinguistics/corpus linguistics databases applications fall short. 


\subsection{Functionality}

This application will be able to perform the necessary functions needed by field linguists. The dashboard will be composed of several widgets. The Data Entry widget will be the primary focus, containing four core fields customary for a gloss format (utterance, morpheme-segmentation, gloss, translation). In addition to these, researchers will be able to add customized fields, such as phonetic transcription or context for an utterance.  Researchers can even upload audio files and link them to the appropriate data. Each data entry will be tagged for the researcher, date of elicitation and consultant's code. 

Furthermore, researchers will be able to add tags for categorization and mark the status of each individual data entry as "Checked" or "To be checked with consultant", which further aids organization and reduces the number of errors that inevitably occur during research. Other functions such as importing data and exporting data into various formats aid efficiency and convenience. 

Another widget will be an Activity Feed View displaying the most recent changes. This widget allows researchers to keep up to date on their team's activity.  The activity feed will display items such as recent additions to the corpus, comments made on data entries, and recent edits. 

Finally, the application will have a powerful search function that will expand into a data list that is contained in its own widget. Data lists can be sorted, saved and can be used for batch operations such as exporting or converting into LaTeX. 

This application integrates the best functions from existing fieldwork database programs, while avoiding many of the shortcomings discussed above. The core features are summarized as follows: 

\begin{itemize}


\item {\bf Modern}

\begin{itemize}
\item { \bf Simple} The system will be designed to replace Word Documents or LaTeX documents which is a very common way field linguists store data because it requires no training, doesn't require a complicated set-up for data categories, and takes no time to add new categories.  

\item {\bf Attractive} The system will have a modern design like many of the popular websites such as Google and be customizable so that the user can include a picture of where they are doing research as a background.

\end{itemize}

\item {\bf Powerful}
\begin{itemize}
\item {\bf Smart.} The application will guess what users do most often, and automate the process for them. Most importantly, the system will have automated predictable glossing information.
\item {\bf  Searchable.} The application will be designed for search as this is one of the most fundamental tasks a language researcher must be able to do. The search will go far beyond traditional string matches and database indexes; it will be able to display data in context.
\end{itemize}

\item{\bf Data-Centric}
\begin{itemize}
\item {\bf Atheoretical.} The application will not include categories or linguistic frameworks or theoretical constructs that must be tied to the data.  The application will allow an analysis to develop organically as data collection proceeds, as opposed to imposing a particular construct upon entry.  Researchers will be able to set and change their own categories for the data whenever they choose to.
\item {\bf Collaborative.} The system will have users and teams, and permissions for corpora. Permissions will ensure that data can be safely shared and edited by multiple users. Moreover, the corpus will be versioned so that users can track changes and revert mistakes.
\item {\bf Sharable.} The application will allow researchers to share their data with anyone interested in their work. The application will be able to import data from ELAN XML, CSV and text file formats, and to export to XML, Latex and WIki formats. These import/export functions will make it easier to exchange data with those who are not using the same application. 
\end{itemize}


\item{\bf Accessible}
\begin{itemize}
\item { \bf Cross-Platform.} The application will be available for any device that runs a browser. Specifically, the application will be installable as a Chrome extension on Mac, Linux, and Windows computers, as well as Android tablets and phones, and will run both online and offline on these devices. The application will run online on iPads and iPhones. 
\item  {\bf Portable.} Touch tablets are one of the easiest tools to carry and use in field; they have a long battery life; they can play videos or show images for the consultant to elicit complicated contexts; and they permit recording audio and video and direct publishing to YouTube and/or other services. Furthermore, Android tablets are particularly easy to program and integrate the microphone directly into the database.
\item {\bf Work offine.} Running a webapp offline will have considerable consequences for how data is stored, how data is retrieved, and how much data can be used while offline. Most browsers have limits on the amount of data a webapp can store offline. By delivering a version of the application in a Chrome extension, which has permission to have unlimited storage, researchers will be able to have a significant portion of their data at their fingertips, regardless of the location. 
\item {\bf Multiple storage} Data will be stored on a CouchDB server, and will be accessible and sharable by multiple users.  The installable Chrome extension version of the application allows individual researchers to store a portion of a corpus, or an entire corpus, on their own devices, enabling offline work and quick search. 
% Audio files will be stored elsewhere (not on a CouchDB server), most likely on researcher's own server, to keep the storage space small in order to provide . The database will contain the information of the location path of the audio files, and when users want to access audio files, they will be redirected to their location. We'll use Iris CouchDB which provides free service if the usage (storage and transaction) is less than 5$/month (http://www.iriscouch.com/service). If we host only text data, it is possible to keep the service free. 
\end{itemize}



\item{\bf Open}
\begin{itemize}
\item {\bf OpenData}. Corpora often contain sensitive information, informant stories and other information which must be kept confidential. Having confidential data in plain text in a corpus forces the entire corpus to be kept confidential. Instead, the system will encrypt confidential data and store the data in the corpus encrypted. To access the plain text the user will have to log in and use a password to decrypt the data. This design has important ramifications for exporting data, and for editing the data outside the application.
\item { \bf OpenSource}. Being OpenSource allows departments to install and customize the database application to tailor specific needs without worry that the company behind the software will disappear or stop maintaining the software. In addition, OpenSourcing the software on GitHub will allow linguists with scripting or programming experience to contribute back to the software to make it more customized to their needs, language typologies, or linguistics research areas.
\item {\bf Unicode}. Encoding problems and losing data should be behind us in the days of unicode. However, many existing fieldlinguistics databases were built in programming languages that did not support unicode, so the unicode support is dangerously fragile.
\end{itemize}

\end{itemize}


\section {Goals \& Objectives}

The principal goal of The FieldLinguists' App is to help language researchers collect and organize linguistic data and to facilitate collaborative research work. The main objectives are to provide: 

\begin{itemize} 
\item A self-explanatory, easy-to-use user interface so that researchers can start using the application at the time the installation is completed.  
\item Both on- and off-line functionality so that the fieldwork is not constrained by the internet accessibility. 
\item Customizable data entry fields to accommodate particular requirements of a research.  
\item Data sharing and integration functions to facilitate collaboration among researchers and between researchers and language consultants. 

\end{itemize} 


Although it is designed primarily for linguists, the application will equally be useful for researchers documenting endangered languages and/or creating dictionaries/grammar books for minority languages, as well as language teachers creating educational materials.  





\section {Budget \& Timeline}

The FieldLinguists' App is composed of eight modules and thus the cost is divided into eight major components. In addition, a separate price is given for software architecture and for 1 year of user support and project growth, which is needed to make a longterm viable and useful tool that field linguists can adopt for their labs or for their field methods courses. The cost is calculated by determining the time in hours and multiplying by \$42-58, the average rate of a software developer in Montr\'eal. 

Since the budget is dependent on how long it takes to complete the modules, it is possible to focus on each module separately, thus reducing the time of completion. 

 

\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline
    Module & Weeks & Price\\ 
\hline
    Software Architecture & 0.5  & \$1,555.20  \\ 
    Collaboration Module & 2.5  & \$5,728.32  \\ 
    Corpus Module & 4.2 & \$9,201.60 \\
     Web Spider Module  & 2.5 & \$2,177.28\\
      Lexicon Module & 2.0 & \$7,340.54 \\ 
     Phonological Search Module & 3.0 & \$2,177.28 \\
    Dictionary Module & 22.3 & \$18,781.63 \\
        Glosser Module & 19.7 & \$17,770.75 \\
Aligner Module & 10.2 & \$ 9,787.39\\ 
User Support & 21.1 & \$30,246.70 \\
TVS and TPQ &  & \$10,833.32 \\
Total &  88 &\$83,176.04\\

\hline
  \end{tabular}
  \caption{Project Summary}
  \label{tab:label}
  \end{center}
\end{table}


\newpage

\subsection{Collaboration Module}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline
Iteration&	Hours&	Technology	\\
\hline
Software Architecture Design&	20&	Software Engineering	\\ 
Collaboration API on central server&	30&	Software Engineering\\ 
Users Model&	15&	Javascript	\\ 
Informant Model&	15&	Javascript	\\ 
Team Model&	15&	Javascript	\\ 
Bot Model&	15&	Javascript	\\ 
User Activity Model&	8&	Javascript	\\ 
Team Feed Widget&	25&	HTML5	\\ 
User list item Widget&	16&	HTML5	\\ 
Team Preferences Widget&	8&	HTML5	\\ 
User Profile Widget&	8&	HTML5	\\ 
User Tests&	30&	Javascript \\ 
Informant Tests&	30&	Javascript	\\ 
Team Tests&	30&	Javascript	\\ 
Android Deployment&	15&	Java	\\ 
Chrome Extension Deployment&	20&	Javascript \\ 
Heroku Deployment&	5&	Integration	\\ 
%\# of weeks with 3 full time personnel&	2.5416666667& \\ 
\hline
  \end{tabular}
  \caption{The Collaboration Module is used to permit collaboration with teams and users. }
  \label{tab:label}
  \end{center}
\end{table}



\subsection{Corpus Module}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline

Iteration&	Hours&	Technology	\\
\hline
Software Architecture Design&	20&	Software Engineering	\\ 
Corpus API on corpus server&	20&	Software Engineering\\ 
Corpus Model&	8&	Javascript	\\ 
Session Model&	8&	Javascript	\\ 
Datum Model&	8&	Javascript	\\ 
Datum status model&	8&	Javascript	\\ 
DataList Model&	8&	Javascript	\\ 
Confidential datum encrypter&	16&	Javascript	\\ 
Audio upload and play logic&	8&	Javascript	\\ 
Corpus DB  implementation on Android&	20&	Java	\\ 
Corpus DB  implementation on Chrome&	20&	Javascript	\\ 
Corpus DB  implementation on Node.js&	20&	Javascript \\ 
Corpus versioning Logic&	25&	Javascript	\\ 
Corpus Preferences Widget&	6&	HTML5	\\ 
Session Preferences Widget&	6&	HTML5	\\ 
Datum Preferences Widget&	20&	HTML5 \\ 
Datum Status Preferences Widget&	16&	HTML5	\\ 
DataList Preferences Widget&	6&  HTML5 \\ 
Corpus sync logic&	10&  Javascript \\ 
Corpus diff Widget (to show before sync)&	10&  HTML5 \\ 
Insert Unicode Character Widget&	10&  HTML5 \\ 
Corpus Details Widget&	6&  HTML5 \\
Session Details Widget&	6&  HTML5 \\  
Datum Details Widget&	20&  Javascript \\ 
DataList Widget&	30&  Javascript \\ 
Global Search logic&	30&  Javascript \\ 
Power Search logic&	80&  Javascript \\ 
Corpus Tests&	5&  Javascript \\ 
Session Tests&	10&  Javascript \\ 
Datum Tests&	10&  Javascript \\ 
Datum Status Tests&	10&  Javascript \\ 
DataList Tests&	20&  Javascript \\ 
Heroku Deployment&	5&  Integration \\ 
\hline
  \end{tabular}
 \caption{The Corpus Module is used to sync, share, edit, tag, categorize and open data. }
  \label{tab:label}
  \end{center}
\end{table}
%# of weeks with 3 full time personnel	4.20833333333333



\newpage
\subsection{Web Spider Module}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline
Iteration&	Hours&	Technology	\\
\hline
Corpus Visualization Widget&	40&	HTML5	\\ 
Web Spider Training Logic&	60&	Java	\\ 
\hline
  \end{tabular}
  \caption{This module is a subportion of the Corpus Module.}
  \label{tab:label}
  \end{center}
\end{table}
%# of weeks with 1 full time personnel	2.5


\newpage
\subsection{Lexicon Module}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline

Iteration&	Hours&	Technology	\\
\hline
Software Architecture Design&	20&	Software Engineering	\\ 
Lexicon API on Lexicon server&	20&	Software Engineering\\ 
Lexicon Model&	6&	Javascript	\\ 
Morpheme Model&	6&	Javascript	\\ 
Allomorph Model&	6&	Javascript	\\ 
Gloss Model&	6&	Javascript	\\ 
Orthography Model&	16&	Javascript	\\ 
Lexicon DB  implementation on Android&	20&	Java	\\ 
Lexicon DB  implementation on Chrome&	20&	Javascript	\\ 
Lexicon DB  implementation on Node.js&	20&	Javascript \\ 
Lexicon versioning Logic&	10&	Javascript	\\ 
Lexicon Preferences Widget&	6&	HTML5	\\ 
Morpheme Tests&	6&	Javascript	\\ 
Allomorph Tests&	6&	Javascript \\ 
Gloss Tests&	6&	Javascript	\\ 
Orthography Tests&	8&  Javascript \\ 
Lexicon Analysis Widget&	10&  HTML5 \\ 
Lexicon sync logic&	10&  Javascript \\ 
Lexicon diff Widget (to show before sync)&	10&  HTML5 \\ 
Lexicon Details Widget&	6&  HTML5 \\
Lexicon Tests&	12&  Javascript \\  
Heroku Deployment&	5&  Integration \\ 
\hline
  \end{tabular}
 \caption{The Lexicon Module is used to house, and read lexicon entries to be used for the glosser.}
  \label{tab:label}
  \end{center}
\end{table}
%# of weeks with 3 full time personnel	1.95833333333333

\newpage
\subsection{Phonological Search Module}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline
Iteration&	Hours&	Technology	\\
\hline
Phonology Ontology for phonological search&	60&	Java	\\ 
Lexicon Visualization Widget&	40&	Javascript	\\ 
Lexicon Editing Widget&	20&	Javascript	\\ 
\hline
  \end{tabular}
 \caption{This module is a subportion of the Lexicon Module.}
  \label{tab:label}
  \end{center}
\end{table}
%# of weeks with 1 full time personnel	3


\newpage
\subsection{Dictionary Module}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline

Iteration&	Hours&	Technology	\\
\hline
Software Architecture Design&	40&	Software Engineering	\\ 
Dictionary API on Lexicon server&	30&	Software Engineering\\ 
Semantic Model&	60&	Javascript	\\ 
Syntactic Model&	60&	Javascript	\\ 
Citation Model&	60&	Javascript	\\ 
Synonyms Model&	60&	Javascript	\\ 
Dictionary DB implementation&	80&	Integration	\\ 
Dictionary Training Logic&	80&	Java	\\ 
Web Spider Training Logic&	100&	Java	\\ 
Dictionary Preferences Widget&	8&	HTML5	\\ 
Dictionary WordNet Analysis Widget&	120&	HTML6	\\ 
Dialect Profile Widget&	8&	HTML5 \\ 
Semantic Tests&	30&	Javascript	\\ 
Syntactic Tests&	30&  Javascript \\ 
Citation Tests&	         30&  Javascript \\ 
Synonyms Tests&	30&  Javascript \\ 
Spider Tests&	30&	Java \\ 
Training Tests&	30&	Java	\\ 
Heroku Deployment&	5&  Integration \\ 
\hline
  \end{tabular}
 \caption{The Dictionary Module is used to share the lexicon in the form of a WordNet/Wiktionary dictionary with the language community as required by some grants.}
  \label{tab:label}
  \end{center}
\end{table}
%# of weeks with 1 full time personnel	22.275


\newpage
\subsection{Glosser Module}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline

Iteration&	Hours&	Technology	\\
\hline
Software Architecture Design&	40&	Software Engineering	\\ 
Glosser API on Lexicon server&	30&	Software Engineering\\ 
Morpheme Model&	15&	Javascript	\\ 
Allomorph Model&	15&	Javascript	\\ 
Gloss Model&	15&	Javascript	\\ 
Orthography Model&	30&	Javascript	\\ 
Glosser DB implementation&	80&	Integration	\\ 
Glosser Prediction Logic&	80&	Java	\\ 
Glosser Machine Learning Logic&	80&	Java	\\ 
Glosser Training Logic&	80&	Java	\\ 
Web Spider Training Logic&	80&	Java	\\ 
Glosser Preferences Widget&	8&	HTML5	\\ 
Morphological Analysis Widget&	40&	HTML6	\\ 
Dialect Profile Widget&	8&	HTML5 \\ 
Morpheme Tests&	30&	Javascript	\\ 
Allomorph Tests&	30&  Javascript \\ 
Gloss Tests&	         30&  Javascript \\ 
Orthography Tests&	30&  Javascript \\ 
Spider Tests&	30&	Java \\ 
Training Tests&	30&	Java	\\ 
Heroku Deployment&	5&  Integration \\ 
\hline
  \end{tabular}
   \caption{The Glosser Module is used to automatically gloss datum, smarter than the standard lexicon.}
  \label{tab:label}
  \end{center}
\end{table}
%# of weeks with 1 full time personnel	19.65


\newpage
\subsection{Aligner Module}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline

Iteration&	Hours&	Technology	\\
\hline
Software Architecture Design&	10&	Software Engineering	\\ 
Aligner API on Lexicon server&	10&	Software Engineering\\ 
Dictionary Model&	15&	Javascript	\\ 
Aligner DB implementation&	80&	Integration	\\ 
Aligner Machine Learning Integration&	80&	Java	\\ 
Aligner Preferences Widget&	8&	HTML5	\\ 
Audio Waveform Visualization logic&	30&	Javascript	\\ 
Audio Spectrogram Visualization logic&	?&	Javascript \\ 
Transcription User Interface&	80&	HTML5	\\ 
TextGrid export&	20&  Javascript \\ 
Dialect Profile Widget&	         8&  HTML5 \\ 
Orthography Tests&	30&  Javascript \\ 
Training Tests&	30&	Java	\\ 
Heroku Deployment&	5&  Integration \\ 
\hline
  \end{tabular}
  \caption{The Aligner Module is used to create TextGrids from the orthography and the audio files, used for prosody and phonetic analysis.}
  \label{tab:label}
  \end{center}
\end{table}
%# of weeks with 1 full time personnel	10.15

\newpage
\subsection{User Support}

\footnotesize
\begin{table}[htbp]
\begin{center}
  \begin{tabular}{ | lcl | }
\hline

Iteration&	Hours&	Technology	\\
\hline
Sample data&	30&	Linguistics	\\ 
Integrate software with sample data&	30&	Javascript\\ 
Screencasts on how to use the app(s)&	24&	Quicktime/YouTube	\\ 
Screencasts on how to modify the code&	40&	Quicktime/YouTube	\\ 
Server maintenance&	20&	Integration	\\ 
Monitor server costs and develop pricing plan&	100&   Business	\\ 
Answer user emails&	250&	Support	\\ 
Read twitter feeds and facebook channels&	100&	Support \\ 
Help IT/developers install and set up \\ the server on their department servers&	50&	Support	\\ 
Upgrade javascript/android libraries&	40&  Javascript \\ 
Amazon EC2 server CPU+Memory+Bandwidth&	         &  Server \\ 
Release new versions&	160&  Javascript \\ 
\hline
  \end{tabular}
 \caption{User Support includes 1 year of product support and project growth. It is needed to make a longterm viable and useful tool that field linguists can adopt for their labs or for their field methods courses.} 
  \label{tab:label}
  \end{center}
\end{table}
%# of weeks with 1 full time personnel	21.1














%%Include in the budget all expenses for the app, including necessary training costs. Mention any co-funding that you are using from other sources. You may want to include a brief narrative of expenses along with a table of individual cost components. 


\normalsize




\section {Evaluation}

The usefulness and effectiveness of the FieldLinguists' App will be evaluated against the following criteria, taken and modified from E-MELD Best Practices in Digital Language Documentation (http://emeld.org/school/what.html) and DataONE Primer on Data Management (http://www.dataone.org/sites/all/documents/DataONE\_BP\_Primer\_020212.pdf). 

\begin{itemize} 

\item Collection: Data are collected and organized with ease. \\ Self-explanatory UI and on- and off-line functionality will make data collection easier.    

\item Quality: The quality of data is assured. \\ Datum status and comments functions enables researchers to check and inspect the data quality.  

\item Description: Data are annotated and categorized according to various terminological conventions.  \\ Data fields and tags accommodate metadata categories that researchers choose to use. The application is not tied to a particular terminological convention. 

\item Integration: Data from disparate sources are combined into one consistent data set. \\ Sync function helps integrate data taken by multiple researchers into one corpus. Import/export functions enables conversion of data format, making data sharing easier.  

\item Analysis: The application organizes data in ways to help data analysis. \\  Customizable data entry fields and data tags, as well as search/list view functions allow researchers to organize data ready for analysis.  

\item Format: Data are intelligible regardless of the types of operating system. \\ The application runs on PC, Linux and Mac OS as a Chrome extension, as well as on Android. Data supports Unicode and exportable to XML files. 

\item Discovery: Data are searchable and discoverable. \\ The application is accessible through general internet search. Within the application, data are discoverable via keyword search. The two-step data discovery process will minimize irrelevant search results.   

\item Access: Data are accessible. \\ Data stored in the application are open to public view in principle, however authorized researchers (authors of data) have control over who can see and/or edit their data. 

\item Citation: Data provide citation information. \\ Data source will be cited with the name of the corpus in fieldlinguists.com (e.g. fieldlinguists.com\#corpus/corpus-name).   
 
\item Preservation: Data are archived in a way that withstands long-term preservation. \\ Data are stored in a host server as JSON files. JSON files are human-readable text files, therefore the information content would not be lost even if the data format becomes obsolete. 

\item Security: Confidential information must be kept from public access. \\ Encryption will ensure that confidential information such as language consultant's identity be hidden from public access unless the consultant decides otherwise. 


\end{itemize} 


%
%Provide information on the metrics that will be used to determine the effectiveness of the project. It has to meet certain data management criteria -- any scientist can discover and use the data over time; it will benefit the scientific community. We are abiding by the best practice standards set by DataOne.   
%
%Example format: 
%
%\exg. n-tuop'ti-m\\
%1-window-POSS.SG\\
%`My window'
%
%



\end{document}


